---
title: "3.3_Calculation_sea_distances.Rmd"
author: "Iris Van Paemel"
date: "2024-06-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.3_Calculation_sea_distances.R
### Input
The input data consists of csv files:
- Output/Species_Location.csv => header consists of "Specieslist", "BelgianCoast", "Getxo", "Koster", "Laeso", "Limfjord", "Plymouth", "Roscoff", "SwedishWestCoast", "TZS", "Vigo"
- Inputs/Coordinates.csv => header consists of "Observatory.ID", "Latitude", "Longitude"

### Output
The output data consists of csv files:
- sea_distances/[species_name]_distancesTo_[Location_name].csv
=> header consists of "x", "year", "month", "country"
- fly_distances/[species_name]_distancesTo_[Location_name].csv
=> header consists of "x", "year", "month", "country"
- OccurrenceData_test/[species_name].csv => header consists of "Longitude", "Latitude", "year", "month", "country"

### script + descriptions

- First install and load the packages
- Set the working directory and load the input data

```{r,  load libraries and data}
##################################################################################
# This script is an alternative script for calculating sea distances and fly distances
##################################################################################

############################################################################################
# LOAD PACKAGES
############################################################################################
library("gdistance")
library("dplyr")
require("geosphere")
require("rgbif")
library("ggplot2")
library("tidyr")
library("worrms")
library("sf")
library("sp")
library("raster")
library("rnaturalearth")
library("rnaturalearthdata")
library("maps")
library("FRK")


############################################################################################
# LOAD DATA
############################################################################################
# Set working directory to directory where the R-script is saved
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# Read a species list
df <- read.csv("Output/Species_Location.csv")
# Read coordinates file
Coordinates <- read.csv("Inputs/Coordinates.csv")

```

RESHAPE DF (line 38 in 3.3_Calculation_sea_distances.R script)

- The pivot_longer function is used to reshape df to long format, all columns in df except Specieslist will be pivoted. They will be treated as value columns
- Only the values that are bigger than 0 will be kept

```{r Reshape df from wide to long format}
############################################################################################
# RESHAPE DF WITH LOCATIONS AND SPECIES FROM WIDE TO LONG FORMAT
############################################################################################

long <- pivot_longer(df, !Specieslist)
# from tidyr package, reshape df from wide to long format
long <- long[long$value > 0, ]
```

FETCH OCCURRENCE DATA (line 50 in 3.3_Calculation_sea_distances.R script)

- Ensure_columns function:
  - uses df and required columns
  - setdiff is used to find which columns are present in required_columns but not in colnames(df)
  - the for loop will check if the column in missing columns doesn't exist, and then makes this
  column and adds NA values to it
- Fetch_data_in_batches function:
  - set start to 0 = start in list of records in GBIF
  - start combined_data dataframe
  - fetch records from GBIF with settings
  - put res$data into the combined_data dataframe
- assign the required columns
- start a for loop that iterates over species names in the long dataframe
  - print the species name
  - put the filename into the variable 'occurrence_coord'
  - check if the file exists, if so, read the file
  - else: start fetching the data by using the function fetch_data_in_batches
  - Initialize an empty list to store processed data frames
  - Loop over each data frame in the list, ensure columns, and select required columns
  - check if temp_df is not NULL and not empty
  - Ensure the required columns are present
  - else: print that the dataframe is empty.
- check if length of processed_data is bigger than zero: batches of data
- merge the lists of different batches together into res_total
- check the sum of NA values in res_total
- else: print an error message
- rename the columns
- remove occurrences where longitude or latitude is NA
- If the amount of rows or res_total = 0 then:
  - check if the directory test_outputs/errors exists
  - if not: create one and write the error file
  - if it does exist: write the error file
- Lastly: write the file for this species with res_total




```{r fetch occurrence data from GBIF}
#####################################################################################
# FIRST CHECK IF FILE WITH OCCURRENCE DATA IN OCCURRENCEDATA DIRECTORY EXISTS
# IF NOT: MAKE ONE AND GET DATA FROM GBIF
# limit in occ_data() function is changeable for personal preference
####################################################################################

# Function to ensure all required columns are present (used in coming for loop)
ensure_columns <- function(df, required_columns) {
  missing_columns <- setdiff(required_columns, colnames(df))
  for (col in missing_columns) {
    if (!col %in% colnames(df)) {
      df[[col]] <- NA
    }
  }
  return(df)
}
fetch_data_in_batches <- function(species_name, basisOfRecord, batch_size = 10000) {
  start <- 0
  combined_data <- data.frame()
    res <- occ_data(scientificName = species_name, 
                    hasCoordinate = TRUE, 
                    limit = batch_size,
                    basisOfRecord = basisOfRecord,
                    continent = "europe")
  combined_data <- rbind(combined_data, res$data)
  Sys.sleep(1)  # Adding a small delay to be polite to the server
  
  return(combined_data)
}

required_columns <- c('decimalLongitude', 'decimalLatitude', 'year', 'month', 'country')

for (species_name in unique(long$Specieslist)){
  print(paste0("species_name: ", species_name))
  # if file exists:        put content into variable res
  # if file doesn't exist: make one, get data, and put into variable res
  occurrence_coord <- paste0("OccurrenceData_test/", species_name, ".csv")
  if (file.exists(occurrence_coord) == TRUE) {
    res <- read.csv(occurrence_coord, header = TRUE)
  } else {
    data_list <- list(
      fetch_data_in_batches(species_name, "Observation"),
      fetch_data_in_batches(species_name, "Machine observation"),
      fetch_data_in_batches(species_name, "Human observation"),
      fetch_data_in_batches(species_name, "Material sample"),
      fetch_data_in_batches(species_name, "Living specimen"),
      fetch_data_in_batches(species_name, "Occurrence"))

    # Initialize an empty list to store processed data frames
    processed_data <- list()
  
    # Loop over each data frame in the list, ensure columns, and select required columns
    for (i in seq_along(data_list)) {
      temp_df <- data_list[[i]]
    
      if (!is.null(temp_df) && nrow(temp_df) > 0) { # check if temp_df is not NULL and not empty
        # Ensure the required columns are present
        temp_df <- ensure_columns(temp_df, required_columns)
        temp_df <- temp_df[, required_columns]
        processed_data[[i]] <- temp_df
        
      } else {
        print(paste0("Data frame ", i, " is empty."))
      }
    }
  }
    
  if (length(processed_data) > 0) {
    res_total <- do.call(rbind, processed_data)
    print(paste0("Total NA values in res_total: ", sum(is.na(res_total))))
  } else {
    print(paste0("No data to combine for species: ", species_name))
  }
  #rename the column names
  colnames(res_total) <- c('Longitude', 'Latitude', 'year', 'month', 'country')
  # Remove occurrences where longitude or latitude is NA
  res_total <- res_total[!is.na(res_total$Latitude) & !is.na(res_total$Longitude),]
  # check if there's no information for a species
  if (nrow(res_total) == 0) {
    error_message <- paste0("No information found on GBIF for ", species_name)
    
    # check if directory with error messages exists, if it doesn't: make one
    if (!dir.exists("test_outputs/errors")){
      dir.create("test_outputs/errors", recursive = TRUE)
      error_file_name <- paste0("test_outputs/errors/error_", species_name, ".csv")
      # error files written to test_outputs/errors/
      write.csv(error_message, file = error_file_name)
      return(FALSE)
    
      # write error file to the directory
    } else {
      error_file_name <- paste0("test_outputs/errors/error_", species_name, ".csv")
      # error files written to test_outputs/errors/
      write.csv(error_message, file = error_file_name)
      return(FALSE)
    }
  }
  print("file has successfully been written")
  write.csv(res_total, file = occurrence_coord)
}
```

FIRST PART OF DISTANCE CALCULATION FUNCTION (line 150 in 3.3_Calculation_sea_distances.R script)

- Iterate over species_name and location_name
- Initialize an empty list to store error messages = error_messages
- Define a helper function to add error messages
- Print species name and location
- Try to read occurrence data in a tryCatch function
  - 

```{r}
############################################################################################
# REVISION: CALCULATE DISTANCES
############################################################################################

# Iterate over species_name and location_name
Calculation_seadistance <- function(species_name, species_location){
  # Initialize an empty list to store error messages
  error_messages <- list()
  
  # Define a helper function to add error messages
  add_error_message <- function(message) {
    error_messages <<- c(error_messages, message)
  }
  # Print species name and location
  print(paste0("species_name: ", species_name))
  print(paste0("species_location: ", species_location))
  
  # Try to read occurrence data
  OccurrenceData <- tryCatch({
    occurrence_coord <- paste0("OccurrenceData_test/", species_name, ".csv")
    read.csv(occurrence_coord, header = TRUE)
  }, error = function(e) {
    add_error_message(paste("Error reading occurrence data for", species_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(OccurrenceData)) return(list(result = NA, error_messages = error_messages))
  
  # Try to get coordinates for ARMS location
  # use grep to get the row from the Coordinates df where location_name is present
  location_row_index <- grep(species_location, Coordinates$Observatory.ID)
  if (length(location_row_index) == 0) {
    add_error_message(paste("Location not found in Coordinates for", species_location))
    return(list(result = NA, error_messages = error_messages))
  }
  # save longitude and latitude for the row that you selected with grep
  longitude <- Coordinates[location_row_index, "Longitude"]
  latitude <- Coordinates[location_row_index, "Latitude"]
  # make a dataframe out of the longitude and latitude called samplelocation
  samplelocation <- data.frame(Latitude = latitude, Longitude = longitude)
  
  # check if OccurrenceData has coordinates
  if (nrow(OccurrenceData) < 1) {
    add_error_message("OccurrenceData has no coordinates")
    return(list(result = NA, error_messages = error_messages))
  }
  
    
```





